# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
setwd("~/Coursera/Reproducible Research/RepData_PeerAssessment1")
library(ggplot2)
library(lattice)
library(plyr)
activity <- read.csv(unz("activity.zip", "activity.csv"))
daily <- aggregate(steps~date, data=activity, FUN="sum")
interval <- aggregate(steps~interval, data=activity, FUN="mean")
```


## What is mean total number of steps taken per day?
```{r fig.height=4}
hist(daily$steps, main="Histogram of Steps Per Day", xlab="Steps Per Day")
stepsMean <- mean(daily$steps)
stepsMedian <- median(daily$steps)
```

The mean number of steps per day is `r stepsMean` and the median number per day is `r stepsMedian`


## What is the average daily activity pattern?

```{r}
plot(interval$interval, interval$steps, type="l", ylab="Steps", xlab="Interval", main="Mean of Steps Taken Per Daily Interval")
maxInterval <- interval[which.max(interval$steps), "interval"]
```
The interval with the largest average number of steps per day is `r maxInterval`

## Imputing missing values

There are a number of records with missing values. The table below shows
the rows with complete cases so the total number of rows
with missing values is under "FALSE".


```{r}
table(complete.cases(activity))
```

Next, create a dataset that imputes values for the missing values by
using the average number of steps per interval.

```{r}
imputed <- ddply(activity, .(interval), function(activity) {activity$steps[is.na(activity$steps)] <- mean(activity$steps, na.rm=TRUE); return(activity)})
dailyImputed <- aggregate(steps~date, data=imputed, FUN="sum")
hist(dailyImputed$steps, main="Histogram of Steps Per Day", xlab="Steps Per Day")
stepsMeanImputed <- mean(dailyImputed$steps)
stepsMedianImputed <- median(dailyImputed$steps)
stepsMeanDiff <- stepsMeanImputed - stepsMean
stepsMedianDiff <- stepsMedianImputed - stepsMedian
```

Using the new imputed dataset, the mean number of steps per day is `r stepsMeanImputed` which differs from the original mean by `r stepsMeanDiff`.  

Using the new imputed dataset, the  median number per day is `r stepsMedianImputed` which differs from the original mean by `r stepsMedianDiff`.  


## Are there differences in activity patterns between weekdays and weekends?
```{r}
imputed$date <- strptime(imputed$date, format="%Y-%m-%d")
imputed$weekday <- weekdays(imputed$date)
imputed$weekdayGroup <- ifelse(imputed$weekday %in% c("Saturday", "Sunday"), "Weekend", "Weekday")
intervalWeekday <- aggregate(steps~interval*weekdayGroup, data=imputed, FUN="mean")
xyplot(intervalWeekday$steps~intervalWeekday$interval|intervalWeekday$weekdayGroup, t="l", xlab="Interval", ylab="Steps", main="Mean of Steps Taken Per Daily Interval With Imputed Data")
```
